MinCall --- MinION end2end convolutional deep learning basecaller

### 2017-07-01 paper/main.tex
Reorganised few stuff

Oxford Nanopore Technology's MinION nanopore sequencing platform~\cite{mikheyev2014first} is the first portable DNA sequencing device. It's small weight, of only 90 grams, low capital cost, and long read length combined with high-throughput, real-time data analysis, and decent accuracy yield promising results in various applications. From clinical application such as monitoring infectious disease outbreaks~\cite{judge2015early}\cite{quick2016real}, characterizing structural variants in cancer\cite{norris2016nanopore} and even full human genome assembly~\cite{jain2017nanopore}.

@@

The exact error rate metric is unreliable since multiple pipeline tools could be the issue. First the sample is prepared, hopefully, uncontaminated and matching reference genome as close as possible then sequenced using the MinION device obtaining raw data. Next, our model (or any other) comes along and basecalls the sequences. To evaluate error rate metric basecalled read is aligned to the reference genome using aligners with their own errors/biases, most commonly used BWA-MEM~\cite{li2013aligning} and Graphmap~\cite{sovic2016fast}.

@@

\section{Related work}
The core of the decoding process is the basecalling step, that is translating the current samples to the nucleotide sequence. Nowadays there are multiple basecalling options, what official and unofficial ones.

@@

To help training process, the raw signal is split into smaller blocks that are used as inputs. For each Metrichor basecalled event is easy to determine the block it falls into using \emph{start} field. Using this information output given by Metrichor can be determined for each block.
To correct errors produced by Metrichor and possibly increase the quality of data, each read is aligned to the reference. This is done using aligner GraphMap~\cite{sovic2016fast} that returns the best position in the genome, hopefully, the part of the genome from which read came from.
Alignment part in the genome is used as a target. Using CIGAR string returned by aligner we can correct Metrichor data and get target output for each block. This process is shown in figure~\ref{fg:data_correction}.

@@

Developed tool was compared with other available basecallers that support R9 chemistry. This includes third-party basically DeepNano and official basecallers by Oxford Nanopore (cloud-based Metrichor and Nanonet).
The fact that ground truth is not known makes evaluation difficult.
Different methods for evaluation were used to get clearer information about each basecaller.

>>>>

Oxford Nanopore Technology's MinION nanopore sequencing platform~\cite{mikheyev2014first} is the first portable DNA sequencing device. It's light, only 90 grams. It's has low capital cost. It's featuring long read length combined with high-throughput, real-time data analysis, and decent accuracy. Therefore it yields promising results in various applications. From clinical application such as monitoring infectious disease outbreaks~\cite{judge2015early}\cite{quick2016real}, characterizing structural variants in cancer\cite{norris2016nanopore} and even full human genome assembly~\cite{jain2017nanopore}.

@@

@@

\subsection{Related work}
The core of the decoding process is the basecalling step, that is translating the current samples to the nucleotide sequence. Nowadays there are multiple basecalling options, what official and unofficial ones.

@@

To help training process, the raw signal is split into smaller blocks that are used as inputs. For each Metrichor basecalled event is easy to determine the block it falls into using \emph{start} field. Using this information output given by Metrichor can be determined for each block.

To correct errors produced by Metrichor and possibly increase the quality of data, each read is aligned to the reference. This is done using aligner GraphMap~\cite{sovic2016fast} that returns the best position in the genome, hopefully, the part of the genome from which read came from.

Alignment part in the genome is used as a target. Using CIGAR string returned by aligner we can correct Metrichor data and get target output for each block. This process is shown in figure~\ref{fg:data_correction}.

@@

Developed tool was compared with other available basecallers that support R9 chemistry. This includes third-party basically DeepNano and official basecallers by Oxford Nanopore (cloud-based Metrichor and Nanonet).

The exact error rate metric is unreliable since multiple pipeline tools could be the issue. First the sample is prepared, hopefully, uncontaminated and matching reference genome as close as possible then sequenced using the MinION device obtaining raw data. Next, our model (or any other) comes along and basecalls the sequences. To evaluate error rate metric basecalled read is aligned to the reference genome using aligners with their own errors/biases. We used standard BWA-MEM~\cite{li2013aligning} and Graphmap~\cite{sovic2016fast}.

The fact that ground truth is not known makes evaluation difficult. Different methods for evaluation were used to get clearer information about each basecaller.
